Doubly-Robust Inference in R using \\texttt{drtmle}

### 2022-06-24 paper/using_drtmle_obstudies.Rmd
add related packages; minor touch-ups

---
output:
  bookdown::pdf_document2:
    template: obs_study_skeleton.tex
bibliography: refs.bib
abs_text: |
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference
  to estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence
  intervals and p-values may be computed based on standard asymptotic results,
  such as the central limit theorem, the delta method, and the nonparametric
  bootstrap. However, in high-dimensional settings maximum likelihood estimation
  often breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
  tests for targeted minimum loss-based estimates of the average treatment
  effect, in addition to several other recently-proposed estimators of the
  average treatment effect.
abs_keywords: |
  causal inference, machine learning, semiparametric estimation, treatment
  effects, multiple robustness, targeted minimum loss estimation
---

@@

In many fields, researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure, or a policy intervention.
Often, the treatment cannot be randomized due to ethical or logistical reasons.
This has sparked great interest in developing statistical methodology to
estimate population-level effects from observational data. Estimation of these
effects often requires accounting for putative confounding factors, that is,
factors related to whether a participant receives treatment and to the
participant's outcome. In many settings, researchers measure many potential
confounders, for example using administrative databases or genetic sequencing
technology. Due to the curse of dimensionality, common statistical estimation
techniques are not feasible for such high-dimensional data without restrictive
modeling assumptions. When these assumptions are violated, estimates of the
population-level effect of a treatment may have large bias.

This has led to increased interest in adopting adaptive estimation techniques
from the machine learning literature to control for high-dimensional confounders
when estimating treatment effects. However, facilitating proper inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, the use of standard causal
inference techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa; @robins:2000:epi] is challenging when
utilizing adaptive estimators. In particular, with rare exception, if
cross-validation is used to select tuning parameters for the nuisance
parameters, estimates of the population-level effect will generally be too
biased to achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing*
must be used to select tuning parameters to control the bias
[@undersmoothed_hal]. However, this can prove quite challenging in practice and
few guidelines are available.

A more natural framework for integrating adaptive methods into nuisance
regression estimation are doubly robust approaches, including augmented inverse
probability of treatment weighted (AIPTW) estimation [@robins:1994:jasa] and
targeted minimum loss-based (TML) estimation [@vdlrubin:2006:ijb]. The
asymptotics of these estimators are characterized in part by the convergence
rates of nuisance parameter estimates with respect to a proper norm. This fact
makes the use of cross-validation more appealing for selection of tuning
parameters for estimators of nuisance parameters, thereby providing a more
straightforward means of obtaining an estimate of the effect of interest that
enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield
closed form standard error estimates, which can serve as the basis for
constructing confidence intervals and hypothesis tests.

As an intermediate step, the AIPTW and TML estimators require estimation of two
key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and TML
estimators a natural appeal: inconsistency in the estimation of one nuisance
parameter may be mitigated by the consistent estimation of the other. However,
recent work has shown that the double robustness property does not necessarily
extend to the inferential properties of these estimators when adaptive
estimators of the PS and/or OR are used
[@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb proposed new
TML estimators that have an asymptotic normal sampling distribution with
estimable variance under consistent estimation of *either* the PS *or* the OR,
paving the way for inference that is doubly-robust. @benkeser:2017:biometrika
proposed modifications of these estimators with similar robustness properties
and illustrated their practical performance.

The two proposed procedures that yield doubly-robust inference are quite
involved, notably requiring iterative estimation of additional, complex nuisance
parameters. The \texttt{drtmle} package was motivated by the need for a
user-friendly tool to implement these methods. The package implements the TML
estimators of population-level effects proposed in @vdl:2014:ijb and
@benkeser:2017:biometrika, as well as several extensions including
cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
\texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
existing IPTW implementations with respect to statistical inference. In
particular, the estimator allows an adaptive estimator of the PS to be used in
construction of the IPTW estimator, while yielding valid statistical inference
about population-level effects. Both the TML and IPTW estimators implemented in
the \texttt{drtmle} package are capable of estimating population-level effects for
discrete-valued treatments; moreover, the package provides convenient utilities
for constructing confidence intervals and hypothesis tests about contrasts of
the mean outcome under different levels of treatment. This document explains
some of the key concepts underlying doubly-robust inference and illustrates
usage of the \texttt{drtmle} package.

@@

An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
covariates is used so that this estimator can be equivalently expressed
\begin{equation*}
  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
\end{equation*}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (1) can be equivalently written as
\begin{align}
  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
          \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation (2)
motivates the estimator
\begin{equation*}
  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap.
If adaptive approaches are used to estimate the PS, inference may not be
available without further modification of the propensity score estimator, as we
discuss below.

## Doubly-robust estimators

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a basis for statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits double-robustness in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
limiting normal distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss-based estimation
(TMLE). TMLE is a general methodology that may be used to modify an estimator
of a regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly-robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (4). In addition to possessing the desirable asymptotic
properties, TML estimators have been shown to outperform AIPTW estimators in
finite samples [@porter:2011:ijb]. We refer readers to @tlbook:2011 for
more about TMLE.

# Doubly-robust inference

@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. @vdl:2014:ijb derived
modified TML estimators that are doubly-robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly-robust variance estimators that may be used to
construct doubly-robust confidence intervals and hypothesis tests.
@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
estimators and demonstrated the practical performance of these estimators via
simulation.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (4) in addition to two additional
equations. These equations involve additional regression parameters that we
refer to as the reduced outcome regression (R-OR) and the reduced propensity
score (R-PS), defined respectively as
\begin{align}
  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly-robust not
only with respect to consistency, but also with respect to its limiting
distribution. The author additionally proposed a TMLE algorithm to map initial
estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
equations.

@@

>>>>

---
output:
  bookdown::pdf_document2:
    template: obs_study_skeleton.tex
    latex_engine: xelatex
bibliography: refs.bib
abs_text: |
  Inverse probability of treatment weighted estimators and doubly robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings, maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators. However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one nuisance estimator is inconsistent.
  \texttt{drtmle} implements doubly robust confidence intervals and hypothesis
  tests for targeted minimum loss estimates of the average treatment effect, in
  addition to several other recently proposed estimators of the average
  treatment effect.
abs_keywords: |
  causal inference, machine learning, semiparametric estimation, treatment
  effects, multiple robustness, targeted minimum loss estimation
---

@@

In many fields, researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure, or a policy intervention.
Often, the treatment cannot be randomized due to ethical or logistical reasons.
This has sparked great interest in developing statistical methodology to
estimate population-level effects from observational data. Estimation of these
effects often requires accounting for putative confounding factors -- that is,
factors related to whether a participant receives treatment and to the
participant's outcome. In many settings, researchers measure many potential
confounders, for example using administrative databases or genetic sequencing
technology. Due to the curse of dimensionality, common statistical estimation
techniques are not feasible for such high-dimensional data without restrictive
modeling assumptions. When these assumptions are violated, estimates of the
population-level effect of a treatment can suffer from a large degree of bias.

This has led to increased interest in adopting adaptive estimation techniques
from the machine learning literature to control for high-dimensional confounders
when estimating treatment effects. However, facilitating proper statistical
inference (i.e.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, the use of standard causal
inference techniques, including both G-computation estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa; @robins:2000:epi] is challenging when
utilizing adaptive estimators. In particular, with rare exceptions, if
cross-validation is used to select tuning parameters for the nuisance
parameters, estimates of the population-level effect will be too biased to
achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing* must be
used to select tuning parameters to control the bias [@undersmoothed_hal;
@ertefaie2022nonparametric; @hejazi2022efficient]. However, this can prove quite
challenging in practice, and few guidelines are available.

A more natural framework for integrating adaptive methods into nuisance
regression estimation are doubly robust approaches, including augmented inverse
probability of treatment weighted (AIPTW) estimation [@robins:1994:jasa] and
targeted minimum loss (TML) estimation [@vdlrubin:2006:ijb]. The
asymptotics of these estimators are characterized in part by the convergence
rates of nuisance parameter estimates with respect to a proper norm. This fact
makes the use of cross-validation more appealing for selection of tuning
parameters for estimators of nuisance parameters, thereby providing a more
straightforward means of obtaining an estimate of the effect of interest that
enjoys standard $n^{1/2}$ asymptotics. Based on the efficient influence
function, a key quantity in semiparametric theory, these asymptotic results also
yield closed-form standard error estimates, which can serve as the basis for
constructing confidence intervals and hypothesis tests.

As an intermediate step, the AIPTW and TML estimators require estimation of two
key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TML estimators also enjoy a *double robustness* property. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and TML
estimators a natural appeal: Inconsistency in the estimation of one nuisance
parameter may be mitigated by the consistent estimation of the other. However,
recent work has shown that the double robustness property does not extend to the
inferential properties of these estimators when adaptive estimators of the PS
and/or OR are used [@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid
inference requires consistent estimation of *both* the PS *and* the OR.
@vdl:2014:ijb proposed new TML estimators that asymptotically attain a normal
sampling distribution with estimable variance under consistent estimation of
*either* the PS *or* the OR, paving the way for inference that is doubly-robust.
@benkeser:2017:biometrika proposed modifications of these estimators with
similar robustness properties and illustrated their practical performance.

The two proposed procedures that yield doubly-robust inference are quite
involved, notably requiring iterative estimation of additional, complex nuisance
parameters. The \texttt{drtmle} package was motivated by the need for a open
source tool to implement these methods. The package implements the TML
estimators of population-level effects proposed in @vdl:2014:ijb and
@benkeser:2017:biometrika, as well as several extensions including
cross-validated targeted minimum loss estimation (CV-TMLE). Additionally,
\texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
existing IPTW implementations with respect to statistical inference. In
particular, the estimator allows an adaptive estimator of the PS to be used in
construction of the IPTW estimator, while yielding valid statistical inference
about population-level effects. Both the TML and IPTW estimators implemented in
the \texttt{drtmle} package are capable of estimating population-level effects
for discrete-valued treatments; moreover, the package provides convenient
utilities for constructing confidence intervals and hypothesis tests about
contrasts of the mean outcome under different levels of treatment. This
manuscript explains some of the key concepts underlying doubly robust inference
and illustrates usage of the \texttt{drtmle} package for the \texttt{R}
language and environment for statistical computing [@R].

In tandem with growing interest in and usage of doubly robust estimation
approaches, open source software tools implementing these methods have
proliferated. Notably extant software packages include the \texttt{tmle}
\texttt{R} package [@gruber2012tmle], which implements TML estimators of the
effects of static single time point interventions; the \texttt{ltmle} \texttt{R}
package [@lendle2017ltmle], which implements TML estimators of the effects of
multiple time point interventions, subject to time-dependent confounding and
censoring, as well as estimators of the parameters of a longitudinal working
marginal structural model; the \texttt{tmle3} \texttt{R} package
[@coyle2021tmle3-rpkg], which serves as a generalizable engine for implementing
TML estimators within the \texttt{tlverse} ecosystem [@vdl2022targeted],
designed to provide a common and extensible framework for TML estimation of the
effects of static, dynamic, and stochastic single time point interventions; and
the \texttt{doubleml} \texttt{R} and \texttt{Python} packages
[@bach2021doublemlR; @bach2022doublemlPy], which implement estimators of causal
effects within the double machine learning framework [@chernozhukov2017double;
@chernozhukov2018double], reyling upon Neymon orthogonality and sample-splitting
for robustness. From among this non-exhaustive list, \texttt{drtmle} stands out
as, to our knowledge, the only implementation of statistical procedures designed
to attain doubly robust inference [@benkeser:2017:biometrika], uniquely
positioning \texttt{drtmle} as a robust tool for the estimation of treatment
effects in observational data settings.

@@

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\bar{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
  \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
          \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{eqn:iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

## Doubly robust estimators

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators generally lack a valid basis for asymptotic statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits _double robustness_ in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
normal limiting distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
TMLE is a general methodology that may be used to modify an estimator of a
regression function in a manner that ensures that the modified estimator
satisfies user-selected equations. In the present problem, one can show that if
an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
\begin{equation}\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
asymptotic properties as the AIPTW estimator. That is, the estimator will be
doubly robust and, provided both the OR and PS are consistently estimated and
regularity conditions are satisfied, will have a normal limiting distribution.
TMLE provides a means of mapping an initial estimator of the OR into an
estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
desirable asymptotic properties, TML estimators have been shown to outperform
AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
methodology.

# Doubly robust inference

@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
and TML estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. @vdl:2014:ijb derived
modified TML estimators that are doubly robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, this
work provided closed-form, doubly robust variance estimators that may be used to
construct doubly robust confidence intervals and hypothesis tests.
@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
estimators and demonstrated the practical performance of these estimators via
simulation experiments.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
two additional equations. These equations involve additional regression
parameters that we refer to as the reduced outcome regression (R-OR) and the
reduced propensity score (R-PS), defined respectively as
\begin{align}
  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
\end{align}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and depend on neither the
true OR nor the true PS. Thus, these regressions may be consistently estimated
at reasonably fast rates irrespective of the dimension of $W$ and irrespective
of whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
R-PS $g_{r,n}^{\star}$ satisfied
\begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
\label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly robust
not only with respect to consistency, but also with respect to its limiting
distribution. The author additionally proposed a TMLE algorithm to map initial
estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
equations.

@@

# Acknowledgements

NSH is grateful for the support of the National Science Foundation (award no.
DMS 2102840).
